---
format:
    gfm:
        default-image-extension: ".png"
always_allow_html: true
execute:
  cache: true
  freeze: auto
  dir: meta
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    dpi: 1280
    fig.height: 6
    out.width: 100%
---

```{r}
#| include: false
#| echo: false
pak::pak(
    pkg = "serkor1/SLmetrics@v0.1-1", 
    ask = FALSE
    )
```

# :bookmark: Version 0.1-1

## :hammer_and_wrench: General

* **Backend changes:** All pair-wise metrics are moved from [{Rcpp}](https://github.com/RcppCore/Rcpp) to `C++`, this have reduced execution time by half. All pair-wise metrics are now faster.

## :sparkles: Improvements

* **NA-controls:** All pair-wise metrics that doesn't have a `micro`-argument were handling missing values as according to C++ and {Rcpp} internals. See [Issue](https://github.com/serkor1/SLmetrics/issues/8). Thank you @EmilHvitfeldt for pointing this out. This has now been fixed so functions uses an `na.rm`-argument to explicitly control for this. See below,

```{r}
## 1) define actual and
## predicted classes
actual    <- factor(c("no", "yes", "yes"))
predicted <- factor(c(NA, "no", "yes"))

## 2) calculate
## accuracy with
## and without na.rm
SLmetrics::accuracy(
    actual    = actual,
    predicted = predicted,
    na.rm     = TRUE
)

SLmetrics::accuracy(
    actual    = actual,
    predicted = predicted,
    na.rm     = FALSE
)
```

## :bug: Bug-fixes

* The `plot.prROC()`- and `plot.ROC()`-functions now adds a line to the plot when `panels = FALSE`. See Issue https://github.com/serkor1/SLmetrics/issues/9.

```{r}
## 1) define actual classes
## and response probabilities
actual <- factor(
    sample(letters[1:3], size = 50, replace = TRUE)
)

response <- rbeta(
    n = 50,
    shape1 = 20,
    shape2 = 2
)

## 2) define ROC and
## prROC objects
roc_obj <- SLmetrics::ROC(
    actual   = actual,
    response = response 
)

pr_obj <- SLmetrics::prROC(
    actual   = actual,
    response = response 
)
```

```{r}
## set plot grid
par(mfrow = c(1,2))

## plot data
## with panels = FALSE
plot(roc_obj, panels = FALSE)
plot(pr_obj, panels = FALSE)
```