---
format:
    gfm:
        default-image-extension: ".png"
always_allow_html: true
execute:
  cache: true
  freeze: auto
  dir: meta/CHANGELOG
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    dpi: 1280
    fig.height: 6
    out.width: 100%
---

```{r install version (v0.3-0)}
#| include: false
#| echo: false
pak::pak(
    pkg = "serkor1/SLmetrics@v0.3-0", 
    ask = FALSE
    )
```

# Version 0.3-0

## :sparkles: Improvements

## New Feature

* **Relative Root Mean Squared Error:** The function normalizes the Root Mean Squared Error by a factor. There is no official way of normalizing it - and in {SLmetrics} the RMSE can be normalized using three options; mean-, range- and IQR-normalization. It can be used as follows,

```{r Relative Root Mean Squared Error (v0.3-0)}
# 1) define values
actual <- rnorm(1e3)
predicted <- actual + rnorm(1e3)

# 2) calculate Relative Root Mean Squared Error
cat(
  "Mean Relative Root Mean Squared Error", SLmetrics::rrmse(
    actual        = actual,
    predicted     = predicted,
    normalization = 0
  ),
  "Range Relative Root Mean Squared Error", SLmetrics::rrmse(
    actual        = actual,
    predicted     = predicted,
    normalization = 1
  ),
  "IQR Relative Root Mean Squared Error", SLmetrics::rrmse(
    actual        = actual,
    predicted     = predicted,
    normalization = 2
  ),
  sep = "\n"
)
```

* **Log Loss:** Weighted and unweighted Log Loss, with and without normalization. The function can be used as follows,

```{r logloss (v0.3-0)}
# Create factors and response probabilities
actual   <- factor(c("Class A", "Class B", "Class A"))
weights  <- c(0.3,0.9,1) 
response <- matrix(cbind(
    0.2, 0.8,
    0.8, 0.2,
    0.7, 0.3
),nrow = 3, ncol = 2)

cat(
    "Unweighted Log Loss:",
    SLmetrics::logloss(
        actual,
        response
    ),
    "Weighted log Loss:",
    SLmetrics::weighted.logloss(
        actual,
        response,       
        weights
    ),
    sep = "\n"
)
```

* **Weighted Receiver Operator Characteristics:** `weighted.ROC()`, the function calculates the weighted True Positive and False Positive Rates for each threshold.

* **Weighted Precision-Recall Curve:** `weighted.prROC()`, the function calculates the weighted Recall and Precision for each threshold.

## :boom: Breaking Changes

* **Weighted Confusion Matrix:** The `w`-argument in `cmatrix()` has been removed in favor of the more verbose weighted confusion matrix call `weighted.cmatrix()`-function. See below,

Prior to version `0.3-0` the weighted confusion matrix were a part of the `cmatrix()`-function and were called as follows,

```{r confusion matrix (deprecated) (v0.3-0)}
#| eval: false
SLmetrics::cmatrix(
    actual    = actual,
    predicted = predicted,
    w         = weights
)
```

This solution, although simple, were inconsistent with the remaining implementation of weighted metrics in {SLmetrics}. To regain consistency and simplicity the weighted confusion matrix are now retrieved as follows,

```{r confusion matrix (v0.3-0)}
# 1) define factors
actual    <- factor(sample(letters[1:3], 100, replace = TRUE))
predicted <- factor(sample(letters[1:3], 100, replace = TRUE))
weights   <- runif(length(actual))

# 2) without weights
SLmetrics::cmatrix(
    actual    = actual,
    predicted = predicted
)

# 2) with weights
SLmetrics::weighted.cmatrix(
    actual    = actual,
    predicted = predicted,
    w         = weights
)
```

## :bug: Bug-fixes

* **Return named vectors:** The classification metrics when `micro == NULL` were not returning named vectors. This has been fixed. 

