---
format:
    gfm:
        default-image-extension: ".png"
always_allow_html: true
execute:
  cache: true
  freeze: auto
  dir: meta/CHANGELOG
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    dpi: 1280
    fig.height: 6
    out.width: 100%
---

```{r install version (v0.3-1)}
#| include: false
#| echo: false
pak::pak(
    pkg = "serkor1/SLmetrics@v0.3-1", 
    ask = FALSE
    )
```

# :bookmark: Version 0.3-1

## :sparkles: Improvements

* **OpenMP Support (PR https://github.com/serkor1/SLmetrics/pull/40):** {SLmetrics} now supports parallelization through OpenMP. The OpenMP can be utilized as follows:

```{r OpenMP (v0.3-1)}
set.seed(1903)

## 1) probability distribution
## function
rand.sum <- function(n){
    x <- sort(runif(n-1))
    c(x,1) - c(0,x)
  }

## 2) generate probability
## matrix
pk <- t(replicate(
    n = 100,
    expr = rand.sum(1e3)
    )
)

## 3) calulate entropy
## with and without OpenMP
SLmetrics::setUseOpenMP(TRUE)
system.time(SLmetrics::entropy(pk))

SLmetrics::setUseOpenMP(FALSE)
system.time(SLmetrics::entropy(pk))
```

* **Entropy with soft labels (https://github.com/serkor1/SLmetrics/issues/37):** `entropy()`, `cross.entropy()` and `relative.entropy()` have been introduced. These functions are heavily inspired by {scipy}.  The functions can be used as follows:

```{r entropy (v0.3-1)}
## 1) Define actual
## and observed probabilities

## 1.1) actual probabilies
pk <- matrix(
  cbind(1/2, 1/2),
  ncol = 2
)

## 1.2) observed (estimated) probabilites
qk <- matrix(
  cbind(9/10, 1/10), 
  ncol = 2
)

## 2) calculate entropy
cat(
  "Entropy", SLmetrics::entropy(pk),
  "Relative Entropy", SLmetrics::relative.entropy(pk, qk),
  "Cross Entropy", SLmetrics::cross.entropy(pk, qk),
  sep = "\n"
)
```

## :boom: Breaking changes

* **logloss:** The argument `response` have ben renamed to `qk` as in the `entropy()`-family to maintain some degree of consistency.
* **entropy.factor():** The function have been deleted and is no more. This was mainly due to avoid the documentation from being too large. The `logloss()`-function replaces it.

## :bug: Bug-fixes

* **Plot-method in ROC and prROC (https://github.com/serkor1/SLmetrics/issues/36):** Fixed a bug in  `plot.ROC()` and `plot.prROC()` where if `panels = FALSE` additional lines would be added to the plot.

