---
format:
    gfm:
        default-image-extension: ".png"
always_allow_html: true
execute:
  cache: true
  freeze: auto
  dir: meta/CHANGELOG
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    dpi: 1280
    fig.height: 6
    out.width: 100%
---

```{r install version (v0.3-2)}
#| include: false
#| echo: false
pak::pak(
    pkg = "serkor1/SLmetrics@v0.3-2", 
    ask = FALSE
    )
```

# Version 0.3-2

## :rocket: Improvements

* **Regression metrics (See PR https://github.com/serkor1/SLmetrics/pull/64):** All regression metrics have had their back-end optimized and are now 2-10 times faster than prior versions.
* **LAPACK/BLAS Support (https://github.com/serkor1/SLmetrics/pull/65):** Added LAPACK/BLAS support for efficient matrix-operations.
* **OpenMP:** Enabling/disabling OpenMP is now handled on the `R`-side and obeys `suppressMessages()`. See below:

```{r OpenMP off (suppress message) (v0.3-2)}
## suppress OpenMP messages
suppressMessages(
  SLmetrics::openmp.off()
)
```

## :fire: New features

* **Available threads:** The available number of threads can be retrieved using the `openmp.threads()`. See below:

```{r OpenMP threads (v0.3-2)}
## number of available
## threads
SLmetrics::openmp.threads()
```

## :bug: Bug-fixes

* **Diagnostic Odds Ratio:** The `dor()` is now returning a single `<[numeric]>`-value instead of `k` number of identical `<[numeric]>`-values.

## :warning: Breaking Changes

* **OpenMP Interface:** The interface to enabling/disabling OpenMP support has been reworked and has a more natural flow. The new interface is described below:

```{r OpenMP on (v0.3-2)}
## enable OpenMP
SLmetrics::openmp.on()
```

```{r OpenMP off (v0.3-2)}
## disable OpenMP
SLmetrics::openmp.off()
```

To set the number of threads use the `openmp.threads()` as follows:

```{r set OpenMP threads (v0.3-2)}
## set number of threads
SLmetrics::openmp.threads(3)
```