---
format:
    gfm:
        default-image-extension: ".png"
always_allow_html: true
execute:
  cache: true
  freeze: auto
  dir: .meta/CHANGELOG
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    dpi: 1280
    fig.height: 6
    out.width: 100%
---

# :bookmark: Version 0.3-4

This update has been focused on two three things:

1. Optimization of the back-end by using Armadillo instead of Eigen.
2. Streamlining and extending the documentation
3. Making functions more flexible

As an example on the increased flexibility is the introduction of the `estimator`-argument in classification metrics - the new approach enables new additions of aggregation methods as the field evolves. The "old" approach were limited to three values `NULL`, `TRUE` and `FALSE`. Furthermore the function signatures of the generics have been made more flexible - this will enable possible wrapping packages to freely implement argument names off the generic.

## :sparkles: Improvements

* **Armadillo backend:** All functions have been ported to the C++ Armadillo library, and are heavily templated and Object Oriented. The functions are 5-20x faster than before.
* **Streamlined documentation:** All documentation have been reworked, and are now using generic {roxygen2} templates. The new structure of the documentation is focused on shared documentation and therefore equal metrics like `recall` and `sensitivity` are aliased, and referenced differently - as a result there should be less noise in the documentation. The *creating factor* has been removed, and all examples are simplified.
* **Efficient multi-metric evaluation:** The Precision-Recall and Receiver Operator Characteristics functions now accepts an `indices` argument. The indices takes an `integer`-matrix of corresponding to the sorted probabilities column-wise. See below:

```{r}
## Classes and
## seed
set.seed(1903)
classes <- c("Kebab", "Falafel")

## Generate actual classes
## and response probabilities
actual_classes <- factor(
    x = sample(
      x = classes, 
      size = 1e2, 
      replace = TRUE, 
      prob = c(0.7, 0.3)
    )
)

response_probabilities <- ifelse(
    actual_classes == "Kebab", 
    rbeta(sum(actual_classes == "Kebab"), 2, 5), 
    rbeta(sum(actual_classes == "Falafel"), 5, 2)
)

## Construct response
## matrix
probability_matrix <- cbind(
    response_probabilities,
    1 - response_probabilities
)

## Calculate Precision-Recall
stopifnot(
    all.equal(
        target  = SLmetrics::pr.curve(actual_classes, probability_matrix),
        current = SLmetrics::pr.curve(actual_classes, probability_matrix, indices = SLmetrics::preorder(probability_matrix, TRUE))
    )
)
```

Depending on the system and data, there is a 3x gain in speed. This approach is highly efficient for cases where multiple AUC or curves are to be computed as it avoids sorting the same probability matrix more than once.

### :rocket: New features

* **Hamming Loss:** The fraction of the wrong labels to the total number of labels, i.e. , where is the target, is the prediction, and is the "Exclusive, or" operator that returns zero when the target and prediction are identical and one otherwise. The interface to `hammingloss()` is given below:


```{r}
set.seed(1903)

## classes
classes <- c("Kebab", "Falafel")

## actual and
## predicted classes
actual    <- factor(sample(classes, 10, TRUE))
predicted <- factor(sample(classes, 10, TRUE))
w         <- runif(n = 10)

## calculate hamming
## loss (weighted and unweighted)
SLmetrics::hammingloss(
    actual,
    predicted
)

SLmetrics::weighted.hammingloss(
    actual,
    predicted,
    w = w
)
```

* **Tweedie Deviance:** The interface to `tweedie.deviance()` is given below:

```{r}
## Generate actual
## and predicted values
actual_values <- c(1.3, 0.4, 1.2, 1.4, 1.9, 1.0, 1.2)

predicted_values <- c(0.7, 0.5, 1.1, 1.2, 1.8, 1.1, 0.2)

## Evaluate performance
SLmetrics::deviance.tweedie(
   actual_values, 
   predicted_values
)
```

* **Gamma Deviance:** The interface to `gamma.deviance()` is given below:

```{r}
## Generate actual
## and predicted values
actual_values <- c(1.3, 0.4, 1.2, 1.4, 1.9, 1.0, 1.2)

predicted_values <- c(0.7, 0.5, 1.1, 1.2, 1.8, 1.1, 0.2)

## Evaluate performance
SLmetrics::deviance.gamma(
   actual_values, 
   predicted_values
)
```

* **Poisson Deviance:** The interface to `poisson.deviance()` is given below:

```{r}
## Generate actual
## and predicted values
actual_values <- c(1.3, 0.4, 1.2, 1.4, 1.9, 1.0, 1.2)

predicted_values <- c(0.7, 0.5, 1.1, 1.2, 1.8, 1.1, 0.2)

## Evaluate performance
SLmetrics::deviance.poisson(
   actual_values, 
   predicted_values
)
```

## :bug: Bug-fixes

## :boom: Breaking changes

* **Area under the curve:** The new interface is given below:

```{r}
## Generate x and y
## pair
x <- seq(0, pi, length.out = 200)
y <- sin(x)

## 1.1) calculate area
SLmetrics::auc.xy(y = y,  x = x)
```

* **Receiver Operating Characteristics:** The new interface is given below:

```{r}
## define classes
## and response probabilities
actual   <- factor(c("Class A", "Class B", "Class A"))
response <- matrix(cbind(
    0.2, 0.8,
    0.8, 0.2,
    0.7, 0.3
),nrow = 3, ncol = 2)

## receiver operating curve
SLmetrics::roc.curve(
    actual,
    response
)

## area under the receiver operating
## curve
SLmetrics::auc.roc.curve(
    actual,
    response,
    estimator = 0 # 0: class-wise, 1: micro average, 2: macro average 
)
```


* **Precision-Recall Curve:** The new interface is given below:

```{r}
## define classes
## and response probabilities
actual   <- factor(c("Class A", "Class B", "Class A"))
response <- matrix(cbind(
    0.2, 0.8,
    0.8, 0.2,
    0.7, 0.3
),nrow = 3, ncol = 2)

## precision-recall curve
SLmetrics::pr.curve(
    actual,
    response
)

## area under the precision-recall
## curve
SLmetrics::auc.pr.curve(
    actual,
    response,
    estimator = 0 # 0: class-wise, 1: micro average, 2: macro average 
)
```

* **Entropy:** `entropy()` has been renamed to `shannon.entropy()`. The new interface to `shannon.entropy()` is given below:

```{r}
## Observed probabilities
pk <- matrix(
  cbind(1/2, 1/2),
  ncol = 2
)

## Shannon Entropy
SLmetrics::shannon.entropy(pk)
```

* **Aggregation in classification metrics:** The aggregation flag in the classification functions `micro` have been replaced with the `integer`-argument `estimator` which falls back to class-wise evaluation if misspecified. The new interface is given below and is applicable to all functions that has this argument:

```{r}
set.seed(1903)

## classes
classes <- c("Kebab", "Falafel")

## actual and
## predicted classes
actual    <- factor(sample(classes, 10, TRUE))
predicted <- factor(sample(classes, 10, TRUE))

## recall: class-wise
SLmetrics::recall(
    actual,
    predicted,
    estimator = 0
)

## recall: micro-averaged
SLmetrics::recall(
    actual,
    predicted,
    estimator = 1
)

## recall: macro-averaged
SLmetrics::recall(
    actual,
    predicted,
    estimator = 1
)
```