---
format:
  html:
    code-overflow: wrap
execute: 
  cache: true
knitr:
  opts_chunk:
    comment: "#>"
    messages: true
    warning: false
---

```{r}
#| echo: false
#| include: false
library(SLmetrics)
```

# Using {lightgbm} and {SLmetrics} in classification tasks

In this section a gradient boosting machine (GBM) is trained on the [wine quality](https://archive.ics.uci.edu/dataset/186/wine+quality)-dataset, and evaluated using [{SLmetrics}](https://github.com/serkor1/SLmetrics). The gbm trained here is a light gradient boosting machine from [{lightgbm}](https://github.com/microsoft/LightGBM).[^1]


```{r data}
# 1) load data
# from {SLmetrics}
data("wine_quality", package = "SLmetrics")
```


```{r}
#| code-fold: true
#| code-summary: "Data preparation"

# 1.1) define the features
# and outcomes
outcome  <- wine_quality$target$class
features <- wine_quality$features

# 2) split data in training
# and test

# 2.1) set seed for 
# for reproducibility
set.seed(1903)

# 2.2) exttract
# indices with a simple
# 80/20 split
index <- sample(1:nrow(features), size = 0.95 * nrow(features))

# 1.1) extract training
# data and construct
# as lgb.Dataset
train <- features[index,]
dtrain <- lightgbm::lgb.Dataset(
    data  = data.matrix(train),
    label = as.numeric(outcome[index]) - 1
)
# 1.2) extract test
# data
test <- features[-index,]


# 1.2.1) extract actual
# values and constuct
# as.factor for {SLmetrics}
# methods
actual <- as.factor(
    outcome[-index]
)

# 1.2.2) construct as data.matrix
# for predict method
test <- data.matrix(
    test
)
```

## Training the GBM

The GBM will be trained with default parameters, except for the `objective`, `num_class` and `eval`.

```{r}
# 1) define parameters
training_parameters <- list(
    objective     = "multiclass",
    num_class     = length(unique(outcome))
)
```

### Evaluation function

```{r}
# 1) define the custom
# evaluation metric
eval_fbeta <- function(
    dtrain, 
    preds) {

        # 1) extract values
        actual    <- as.factor(dtrain)
        predicted <- lightgbm::get_field(preds, "label")
        value     <- fbeta(
            actual    = actual,
            predicted = predicted,
            beta      = 2,
            # Use micro-averaging to account
            # for class imbalances
            micro     = TRUE
        )

        # 2) construnct output
        # list
        list(
            name          = "fbeta",
            value         = value,
            higher_better = TRUE 
        )
    
}
```

### Training the GBM

```{r}
model <- lightgbm::lgb.train(
    params  = training_parameters,
    data    = dtrain,
    nrounds = 100L,
    eval    = eval_fbeta,
    verbose = -1
)
```

## Evaluation

The predicted classes can be extracted using `predict()` - these values will be converted to `factor` values

```{r forecasts}
# 1) prediction
# from the model
predicted <- as.factor(
    predict(
        model,
        newdata = test,
        type    = "class"
    )
)
```


```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Safe conversion"
predicted <- factor(
     predict(
         model,
         newdata = test,
         type    = "class"
     ),
     labels = levels(outcome),
     levels = seq_along(levels(outcome)) - 1 
)
```

```{r cmatrix}
# 1) construct confusion
# matrix
confusion_matrix <- cmatrix(
    actual = actual,
    predicted = predicted
)

# 2) visualize
plot(
    confusion_matrix
)

# 3) summarize
summary(
    confusion_matrix
)
```


## Receiver Operator Characteristics

```{r response}
# 1) prediction
# from the model
response <- predict(
        model,
        newdata = test
    )
```

The `response` can be passed into the `ROC()`-function,

```{r}
# 1) calculate the reciever
# operator characteristics
roc <- ROC(
    actual   = actual,
    response = response
)

# 2) print the roc
# object
print(roc)
```

The `ROC()`-function returns a `data.frame`-object, with `r nrow(roc)` rows corresponding to the `length` of `response` multiplied with number of classes in the data. The `roc`-object can be plotted as follows,

```{r}
# 1) plot roc
# object
plot(roc)
```


```{r}
roc.auc(
    actual, 
    response
)
```


[^1]: The wine quality dataset comes (almost) ready for analysis. See the [repo](https://github.com/serkor1/SLmetrics) for more details on the data-manipulation steps taken.