---
format:
  html:
    code-overflow: wrap
execute: 
  cache: true
knitr:
  opts_chunk:
    comment: "#>"
    messages: true
    warning: false
---
# OpenMP

[{SLmetrics}](https://github.com/serkor1/SLmetrics) supports parallelization through [OpenMP](https://www.openmp.org/). In this section this functionality is introduced.

## Enabling/Disabling OpenMP

OpenMP is disabled by `default`, but can be enabled as follows:

```{r}
SLmetrics::openmp.on()
```

And disabled as follows:

```{r}
SLmetrics::openmp.off()
```

By `default` all threads are used. To control the amount of threads, see the following code:

```{r}
SLmetrics::openmp.threads(3)
```

To use all threads:

```{r}
SLmetrics::openmp.threads(NULL)
```

## Available threads

The number of available threads are detected automatically, but can also be viewed using `SLmetrics::openmp.threads()` without passing any arguments. See below:

```{r}
SLmetrics::openmp.threads()
```

## Benchmarking OpenMP

To benchmark the performance gain on enabling OpenMP, the same setup as in @sec-benchmarking is used. Below is the `actual` and `predicted` values are generated.

```{r}
#| include: false
#| echo: false
#| code-fold: true
benchmark <- function(
  ..., 
  m = 10) {
  library(magrittr)
  # 1) create list
  # for storing values
  performance <- list()
  
  for (i in 1:m) {

     # 1) run the benchmarks
    results <- bench::mark(
      ...,
      iterations = 10,
      check = FALSE
    )

    # 2) extract values
    # and calculate medians
    performance$time[[i]]  <- setNames(lapply(results$time, mean), results$expression)
    performance$memory[[i]] <- setNames(lapply(results$memory, function(x) { sum(x$bytes, na.rm = TRUE)}), results$expression)
    performance$n_gc[[i]] <- setNames(lapply(results$n_gc, sum), results$expression)

  }

  purrr::pmap_dfr(
  list(performance$time, performance$memory, performance$n_gc), 
  ~{
    tibble::tibble(
      expression = names(..1),
      time = unlist(..1),
      memory = unlist(..2),
      n_gc = unlist(..3)
    )
  }
) %>%
  dplyr::mutate(expression = factor(expression, levels = unique(expression))) %>%
  dplyr::group_by(expression) %>%
  dplyr::filter(dplyr::row_number() > 1) %>%
  dplyr::summarize(
    execution_time = bench::as_bench_time(median(time)),
    memory_usage = bench::as_bench_bytes(median(memory)),
    gc_calls = median(n_gc),
    .groups = "drop"
  )

}
```


```{r}
#| include: false
#| message: false
#| echo: false
# classification function
create_factor <- function(
    k = 3,
    balanced = TRUE,
    n = 1e7) {

  probs <- NULL

  if (!balanced) {

    probs <- rbeta(
      n = k,
      shape1 = 10,
      shape2 = 2
    )

    probs[which.min(probs)] <- 0

    probs <- probs / sum(probs)

  }

  factor(
    x = sample(
      1:k,
      size = n,
      replace = TRUE,
      prob = probs
    ),
    labels = letters[1:k],
    levels = 1:k
  )
}
```

```{r}
# 1) set seed for reproducibility
set.seed(1903)

# 2) create classification
# problem
fct_actual <- create_factor()
fct_predicted <- create_factor()
```


```{r}
#| code-fold: true
#| message: false
#| label: tbl-classification-openmp
#| tbl-cap: Benchmark of computing a 3x3 confusion matrix with OpenMP enabled. Each benchmark is run 10 times with two input vectors of 10 million elements.
SLmetrics::openmp.on()

benchmark(
    `With OpenMP` = SLmetrics::cmatrix(fct_actual, fct_predicted)
)
```


```{r}
#| code-fold: true
#| message: false
#| label: tbl-classification
#| tbl-cap: Benchmark of computing a 3x3 confusion matrix with OpenMP enabled. Each benchmark is run 10 times with two input vectors of 10 million elements.
SLmetrics::openmp.off()

benchmark(
    `Wihtout OpenMP` = SLmetrics::cmatrix(fct_actual, fct_predicted)
)
```

## Key take-aways

Enabling OpenMP support can decrease computation time significantly - but it should *only* be used consciously, and with care, to avoid function calls competing for the same threads. This is especially the case if you are running a, say, neural network in parallel.